---
title: "Untitled"
author: "Julia Esteve"
date: "2026-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Exercise 1. Continent-wide life expectancy
```{r, message=FALSE, warning=FALSE}
library(gapminder)
library(dplyr)
gapminder
```

***1. Compute the mean life expectancy and GDP per capita for each (continent, year)***
```{r, message=FALSE, warning=FALSE}
tab = group_by(gapminder, continent, year) |> summarise(mean_lifeExp= mean(lifeExp, na.rm= TRUE), mean_gdpPercap= mean(gdpPercap, na.rm=TRUE))
tab
```

***2. Do an animated scatterplot of continent life expectancy vs. GDP (transition_time(year))***

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(gganimate)
animation <- ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color=continent)) + geom_point( size = 2) + labs(x= 'GDP per Capita', Y='Life expectancy', subtitle = 'Year={frame_time}') + transition_time(year)
animation
```

## Exercise 2. Inequality of GDP/capita across countries over time
***1. For each year, compute the inter-quartile range of the GDP/capita across countries***
```{r, message=FALSE, warning=FALSE}
IQR_per_year = group_by(gapminder, year) |> summarise(gdpPercap_IQR=IQR(gdpPercap, na.rm = TRUE))
IQR_per_year
```

***2. Do a scatterplot of IQR vs. year. Add lines to the plot***
```{r, message=FALSE, warning=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
